### 2020-07-13 ISSTA/concepts.tex
shrinking agian

We define two basic types of determinism, shown in Figure
\ref{fig:types}:  \emph{horizontal} and \emph{vertical} determinism.  In
horizontal determinism a software system reliably
produces the same behavior given the same steps of a test: the
behavior of multiple executions that are ``the same'' in terms of
inputs/actions can be aligned and checked for equality.  In
vertical determinism, rather than behavior across executions, we are
interested in behavior within an execution, where repeating the same
step of a test twice should result in the same behavior.  Vertical determinism is both rarer and more specialized than
horizontal determinism.

@@

Many \emph{idempotent} operations fall into this category.
Consider adding an element to an AVL tree implementing a set, not a
multiset.  Assume the method call returns the parent node of the
element, whether it was added or was already in the tree.  Calling
this method any number of times in a row should always return the
same value.  One approach, then, would be to identify
idempotent operations 
automatically retry all such operations, checking that the
result is unchanged.  Identifying idempotent operations, however, imposes a specification burden.

@@

In order to implement a practical nondeterminism-detection tool, it is
important to consider the real-world sources of nondeterminism.  
\begin{comment}
Most of these
sources can, in principle, be detected simply by re-running a test
twice and comparing results, either at all visible values or at the
final state.  For example, if some library call uses a random number
generator (perhaps in a stochastic algorithm), the returned value should be
different on a second run.  Timing dependencies may change on a second
execution, especially if a delay is added between steps of the test (a
more principled version of the unfortunate tradition of testing
concurrent code by adding ad-hoc {\tt sleep} statements).
\end{comment}
Most nondeterministic behavior in tests is probably attributable to simple
effects that can vary with a repeated execution, in the same
(approximate) enviroment, of the same test.  There are two broad classes of
nondeterminism that can usually be detected by
executing the same test twice in succession, without attempting
to control any other factors.  {\bf External Environment} nondeterminism arises
  when a test's behavior depends on factors outside the program under
  test, which are subject to uncontrolled variance.  Calling {\tt
    random()} in Python is an obvious example. Similarly, calls to {\tt time}
  functions  will usually return different values on different runs,
  and even if only elapsed time is used, due to system load, I/O, and
  other external factors.  {\bf Concurrency} is a second common cause for
  nondeterminism, when an algorithm is implemented using
  concurrency, in the (incorrect) belief that the result of the
  algorithm remains deterministic.  E.g., a multi-threaded
  implementation of a search algorithm, that returns the location of
  an item in an unsorted list; in the presence of duplicate items,
  thread scheduling may change the index returned.  Lam et al. \cite{idflakies} note that flaky tests not due to order
dependence are generally due to ``concurrency, timeouts,
network/IO, etc.'' a description that is well accounted for by these two
primary sources.

\begin{comment}
\begin{itemize}
\item {\bf External Environment:} This class of nondeterminism arises
  when a test's behavior depends on factors outside the program under
  test, which are subject to uncontrolled variance.  Calling {\tt
    random()} in Python is an obvious example. Similarly, calls to {\tt time}
  functions  will usually return different values on different runs,
  and even if only elapsed time is used, due to system load, I/O, and
  other external factors.
  %Finally, in
  %distributed system testing, the environment is often remote servers
  %(or clients) that may be unavailable, slower than expected, or
  %subject to faulty behavior for reasons not under the control of the
  %test.
\item {\bf Concurrency:} A second common cause for
  nondeterminism in tests is when an algorithm is implemented using
  concurrency, in the (incorrect) belief that the result of the
  algorithm remains deterministic.  E.g., a multi-threaded
  implementation of a search algorithm, that returns the location of
  an item in an unsorted list; in the presence of duplicate items, thread scheduling may change the index returned.
  %If the list may contain duplicates,
  %and a developer has not accounted for this possibility, the
  %algorithm may return different locations of an item
  %depending on the scheduling of threads.
  %Arguably, this can be
  %subsumed under the first class, with the behavior of the
  %processor/scheduler as the ``enviromental'' factor; relying on
  %scheduling is, in a sense, equivalent to calling {\tt random} or
  %{\tt time}.
\end{itemize}
\end{comment}

@@

Another example of process-based nondeterminism was discovered by numerous Python developers when Python
version 3.3 introduced automatic random salting of hashes on a
per-process basis,
in order to mitigate hash-based denial of service attacks
\cite{denial}.  Until version 3.6 this not only resulted in changes in
exact hash values, but in the order of iteration on
dictionaries; in Python 3.6, the default dictionary implementation became an ordered dictionary with consistent iteration, though actual hashes remained nondeterministic.  While few programs relied on exact hash values, many
relied on a
predictable order for dictionary iteration.  Testing
Python code for nondeterminism based on hash seed requires
running in a new process:  Python does not allow changing the salt.

%Other, less common, effects tied to process also exist.  A few tests
%may somehow depend on their actual PID (Process ID) or at least the
%parent process' ID.  Process change increases the likelihood that on a multi-core
%system a test will execute on a different CPU than in the first
%execution, which can have many subtle effects, mostly related to
%changed timing.

@@

A set of $M$ repeated false positives with
$N = \frac{K}{M}$ samples each is much less likely than a false positive
with $K$ samples; so long as we accept the resulting bias in favor of
false negatives, we can therefore produce a reduced test with a
desired $P(\mathit{fail})$ much more cheaply:  the use of replications not only
means we only pay the full sampling price on rare occasions, but a
desired accuracy for $P$ can be obtained with a much smaller value $M
\times N$ than a non-dynamically-sampled $N$.  For example, if $\mathit{PRED}$
is $P(\mathit{fail}) \geq 0.5$, and the true probability for a candidate test
is 0.25, using $N=8$ will give a false positive rate of over 10\%.
Using 4 replications on just 2 samples ($M=4;N=2$) yields a false
positive rate of about 3.7\%, yet requires almost 60\% fewer
executions of the test.  The probability calculations for such comparisons are
relatively simple, but in real testing are
usually not very useful, since the true
probability distributions of test behaviors vary widely and
dynamically during the
delta-debugging process; gathering experimental data during a trial
delta-debugging run and tuning $N$ and $M$ to yield desired
results is  more effective.  In fact, by tuning $M$, any degree of confidence can be achieved, with
the basic tradeoff being between finding a test with the desired
probabilities, and the speed and effectiveness of reduction.
Because increasing $M$ makes false negatives more likely, larger $M$
will usually result in less-than-optimal reduction of the original
test.

To make the basic concepts, more clear, Figures
\ref{fig:p1}-\ref{fig:p3} graphically show the interplay of
delta-debugging and nondeterministic predicates for a simple example.
In the example, tests consist of a sequence of operations that behave
nondeterministically with (independent) probabilities of 0.01, 0.05,
and 0.10, respectively.  Figure \ref{fig:p1} shows one run of
delta-debugging on a test of length 500.  In
the course of reducing the test length to a test with only two
operations, the delta-debugging algorithm also reduces the probability
of nondeterminism from almost 100\% to about 20\%.  If we use a predicate that
``forces'' the test to behave nondeterministically at least half the
time, by sampling the predicate value 100 times and only returning
true when at least 50 of the evaluations report true, we see the
behavior in Figure \ref{fig:p2}:  the final test is slightly longer,
but still falls well short of our target of exhibiting nondeterminism
50\% of the time.  Finally, Figure \ref{fig:p3} shows what happens if
we use the same target of 50\% nondeterminism, but use only 10
samples, with 10 replications ($N$ = 10, $M$ = 10):  the test is not
much longer than in Figure \ref{fig:p2}, but the probability of
nondeterminism is above our target value, close to 60\% (and, as a bonus, many fewer
test executions are required on average for each check of the predicate).

>>>>

We define two basic types of determinism, shown in Figure
\ref{fig:types}:  \emph{horizontal} and \emph{vertical} determinism.  In
horizontal determinism a software system reliably
produces the same behavior given the same steps of a test: the
behavior of multiple executions that are ``the same'' in terms of
inputs/actions can be aligned and checked for equality.  In
vertical determinism, rather than behavior across executions, we are
interested in behavior within an execution, where repeating the same
step of a test twice should result in the same behavior.

@@

Many \emph{idempotent} operations fall into this category.
Consider adding an element to an AVL tree implementing a set, not a
multiset.  Assume the method call returns the parent node of the
element, whether it was added or was already in the tree.  Calling
this method any number of times in a row should always return the
same value.  One approach, then, would be to identify
idempotent operations 
automatically retry all such operations, checking that the
result is unchanged.  Identifying idempotent operations does impose a specification burden.

@@

In order to implement a practical nondeterminism-detection tool, it is
important to consider the real-world sources of nondeterminism.  
\begin{comment}
Most of these
sources can, in principle, be detected simply by re-running a test
twice and comparing results, either at all visible values or at the
final state.  For example, if some library call uses a random number
generator (perhaps in a stochastic algorithm), the returned value should be
different on a second run.  Timing dependencies may change on a second
execution, especially if a delay is added between steps of the test (a
more principled version of the unfortunate tradition of testing
concurrent code by adding ad-hoc {\tt sleep} statements).
\end{comment}
Most nondeterministic behavior in tests is probably attributable to simple
effects that can vary with a repeated execution, in the same
(approximate) enviroment, of the same test.  There are two broad classes of
nondeterminism that can usually be detected by
executing the same test twice in succession, without attempting
to control any other factors.  {\bf External Environment} nondeterminism arises
  when a test's behavior depends on factors outside the program under
  test, which are subject to uncontrolled variance.  Calls to {\tt
    time} or {\tt random} are obvious examples.  {\bf Concurrency} is a second common cause.  E.g., a multi-threaded
  implementation of a search algorithm, that returns the location of
  an item in an unsorted list; in the presence of duplicate items,
  thread scheduling may change the index returned.  
\begin{comment}
\begin{itemize}
\item {\bf External Environment:} This class of nondeterminism arises
  when a test's behavior depends on factors outside the program under
  test, which are subject to uncontrolled variance.  Calling {\tt
    random()} in Python is an obvious example. Similarly, calls to {\tt time}
  functions  will usually return different values on different runs,
  and even if only elapsed time is used, due to system load, I/O, and
  other external factors.
  %Finally, in
  %distributed system testing, the environment is often remote servers
  %(or clients) that may be unavailable, slower than expected, or
  %subject to faulty behavior for reasons not under the control of the
  %test.
\item {\bf Concurrency:} A second common cause for
  nondeterminism in tests is when an algorithm is implemented using
  concurrency, in the (incorrect) belief that the result of the
  algorithm remains deterministic.  E.g., a multi-threaded
  implementation of a search algorithm, that returns the location of
  an item in an unsorted list; in the presence of duplicate items, thread scheduling may change the index returned.
  %If the list may contain duplicates,
  %and a developer has not accounted for this possibility, the
  %algorithm may return different locations of an item
  %depending on the scheduling of threads.
  %Arguably, this can be
  %subsumed under the first class, with the behavior of the
  %processor/scheduler as the ``enviromental'' factor; relying on
  %scheduling is, in a sense, equivalent to calling {\tt random} or
  %{\tt time}.
\end{itemize}
\end{comment}

@@

Another example of process-based nondeterminism arose when Python
version 3.3 introduced automatic random salting of hashes on a
per-process basis,
in order to mitigate hash-based denial of service attacks
\cite{denial}.  Until version 3.6 this not only resulted in changes in
exact hash values, but in the order of iteration on
dictionaries.  Many Python programs
relied on a
predictable ordering.  
%Other, less common, effects tied to process also exist.  A few tests
%may somehow depend on their actual PID (Process ID) or at least the
%parent process' ID.  Process change increases the likelihood that on a multi-core
%system a test will execute on a different CPU than in the first
%execution, which can have many subtle effects, mostly related to
%changed timing.

@@

A set of $M$ repeated false positives with
$N = \frac{K}{M}$ samples each is much less likely than a false positive
with $K$ samples; so long as we accept the resulting bias in favor of
false negatives, we can therefore produce a reduced test with a
desired $P(\mathit{fail})$ much more cheaply, and a
desired accuracy for $P$ can be obtained with a much smaller value $M
\times N$ than a non-dynamically-sampled $N$.

\begin{comment}For example, if $\mathit{PRED}$
is $P(\mathit{fail}) \geq 0.5$, and the true probability for a candidate test
is 0.25, using $N=8$ will give a false positive rate of over 10\%.
Using 4 replications on just 2 samples ($M=4;N=2$) yields a false
positive rate of about 3.7\%, yet requires almost 60\% fewer
executions of the test.  The probability calculations for such comparisons are
relatively simple, but in real testing are
usually not very useful, since the true
probability distributions of test behaviors vary widely and
dynamically during the
delta-debugging process; gathering experimental data during a trial
delta-debugging run and tuning $N$ and $M$ to yield desired
results is  more effective.  In fact, by tuning $M$, any degree of confidence can be achieved, with
the basic tradeoff being between finding a test with the desired
probabilities, and the speed and effectiveness of reduction.
Because increasing $M$ makes false negatives more likely, larger $M$
will usually result in less-than-optimal reduction of the original
test.
\end{comment}

To make the basic concepts, more clear, Figures
\ref{fig:p1}-\ref{fig:p3} graphically show the interplay of
delta-debugging and nondeterministic predicates for a simple example.
In the example, tests consist of a sequence of operations that behave
nondeterministically with (independent) probabilities of 0.01, 0.05,
and 0.10, respectively.  Figure \ref{fig:p1} shows one run of
delta-debugging on a test of length 500.  In
the course of reducing the test length to a test with only two
operations, the delta-debugging algorithm also reduces the probability
of nondeterminism from almost 100\% to about 20\%.  If we use a predicate that
``forces'' the test to behave nondeterministically at least half the
time, by sampling the predicate value 100 times and only returning
true when at least 50 of the evaluations report true, we see the
behavior in Figure \ref{fig:p2}:  the final test is slightly longer,
but still falls well short of our target of exhibiting nondeterminism
50\% of the time.  Finally, Figure \ref{fig:p3} shows what happens if
we use the same target of 50\% nondeterminism, but use only 10
samples, with 10 replications ($N$ = 10, $M$ = 10):  the test is not
much longer than in Figure \ref{fig:p2}, but the probability of
nondeterminism is above our target value, close to 60\% (and, as a bonus, many fewer
test executions were performed).

### 2020-07-13 ISSTA/experiments.tex
paper
shrinking agian

Reducing a lengthy test showing nondeterminism of the swapsort usually takes less than two
minutes (out of 10 trials, only one took four minutes).  The resulting test will be usually be very compact, e.g.
reducing the original test from 56 to 8 steps.

@@

Using the original TSTL harness, just over 20\% of all {\tt redis-py}
regression tests (of length 200) generated were flaky.  Using the
nondeterminism checker to reduce flaky tests to a minimal set allows
us to set up an overnight run that will, in 12 hours, produce
a set of minimal nondeterministic tests exposing the root sources of
nondeterminism/flakiness: we set the harness to repeatedly run, then
produce a regression test.  If the test is flaky (exhibits
nondeterminism in result, over a set of 3 runs), we use the TSTL reducer to shrink it to a
normaliized \cite{onetest} test.  At the end of the testing period,
all distinct tests exhibiting nondeterminism are reported to the
user.  The process reduces the large number of different tests
exhibiting flakiness to a small set of tests, each short, showing a
different cause of nondeterminism.  
There is some duplication of causes, but the tests are sufficient for
identifying all nondeterminism sources.

This process is unlikely to truly require 12
hours, of course; our point is that it is completely automated and
can be given time proportional to the need to find even
low-probability sources of nondeterminism.  The actual amount of time given
may be smaller, or, for that matter, larger (leaving an important
library to self-document nondeterminism over a weekend is hardly a bad
practice).  In this case, the probability of flakiness
is not important, just the possibility, so we let the reducer
create a minimal test with some non-zero (but possibly very small)
observed probability of nondeterminism.  Our 12 hour run identified
all sources of nondeterminism successfully, according to a series of
48 hour runs performed on the version of the harness we modified to
remove nondeterminism, which produced no flaky tests.  We determined that the mean time to find all
the sources of nondeterminism and produce minimal tests is 3.2
hours, which is not trivial, but also, for this purpose, not
excessive.  Our answer to {\bf RQ1} is thus a strong affirmative.

@@

We removed 11 calls from the {\tt redis-py} harness, and  modified 2
calls, after
identification of sources of nondeterminism.  After making these changes, no flakiness was observed in a sample of over 2,000 length
200 tests.   Moreover, because the removals were limited to actually
observed sources of flakiness, the code coverage loss was minimal.
Mean branch coverage, for regression tests of length 200, was reduced
by less than 5 branches (a less than 1\% decrease).  Obviously, it is
impossible to test the removed calls, but the overall coverage loss is
both minimal and known, and can be made up for using specially-crafted
tests (for instance, wrapping the problematic calls in a way that does
not check the values, or adding a delay after an expiration to allow
the data to expire).  As a price to pay for the ability to produce
fast-executing high-coverage full regression tests (not just tests for
crashes and unexpected exceptions), this seems acceptable.  AFL stability was 56.26\% with the harness allowing 
nondeterminism, but rose to 98.32\% using the harness with 
nondeterministic operations removed.

Removing sources of expected nondeterminism also makes
it possible to aggressively test {\tt redis-py} and Redis for
unexpected nondeterminism arising from actual bugs.  The overhead for such
determinism checks, with no delay between operations, is only
20\%, much lower than the expected cost of running each test twice,
answering {\bf RQ2} also in the affirmative.  This is because choosing the actions in a test (and determining which
actions are enabled at each step) consumes a large part of the test
generator's time with a complex library like {\tt redis-py}; running the test again and checking equality is
relatively inexpensive.

As to {\bf RQ3}, Figures \ref{fig:r1}-\ref{fig:r3} show delta-debugging of a typical
{\tt redis-py} nondeterministic test, originally with 287 steps.  The initial test behaves
nondeterministically about half the time.  Figure \ref{fig:r1} shows
that this is a non-monotonic reduction problem, where removing steps
can either increase or decrease the probability of nondeterminism.  At
first, unmodified delta-debugging actually improves the probability of
failure, but eventually it produces a test of length 16 that only behaves
nondeterministically 24\% of the time.  The reduction takes only 38
seconds.  For the purpose of identifying commands leading to
nondeterminism, this is acceptable.  However, if we were actually
debugging a complex nondeterminism bug in Redis itself, we might want
a more reliably nondeterministic test.  Using the same parameters as
in Section \ref{sec:pubbias}, we see the same pattern.  Simply making
a predicate that ``forces'' the probability to remain high, with a
large number of samples, does not work (Figure \ref{fig:r3}) and
requires over 2,000 seconds to produce a test with an even worse
probability of nondeterminism.  Using 10 samples and 10 replications,
on the other hand, actually improves the probability of
nondeterministic behavior, and (in this case) even produces a smaller test (only 14 steps) in just
over 10 minutes.

\subsubsection{Visible Value vs. Final State Nondeterminism}

We also used {\tt redis-py} to investigate the tradeoff between
overhead and ability to detect nondeterminism when using the two
proposed approaches to horizontal determinism.  We produced 300 tests
of length 100, using the known-nondeterministic version of the {\tt
  redis-py} harness.  We then used {\tt nondeterministic} and {\tt
  stepNondeterministic} calls, both with a delay of 0.005 seconds and
a single additonal execution of the test, to
check these tests for nondeterministic behavior.  Final state
nondeterminism actually detected one instance (but not root cause) of nondeterminism that
visible value nondeterminism failed to detect (21 detections
vs. 20). In general, final state
nondeterminism is less able to detect nondeterminism, but in this
instance the difference is less important than the nondeterminism of
whether a particular test will exhibit nondeterministic behavior in a
single run.  This difference is obviously not statistically
significant; for the nondeterminism in {\tt redis-py}, with these
parameters, the two approaches cannot be statistically distinguished
as to effectiveness in detecting nondeterminism.
Final state nondeterminism was also faster; visible value
nondeterminism took, on average, 0.48\% longer to check for
nondeterminism on the 300 tests, 1.199 mean seconds vs. 1.205 mean seconds.  The difference was significant, with
$p$-value $< 1.19 \times 10^{-23}$ by a paired Wilcoxon test.

\begin{comment}
These results, of course, depend on the parameters of the check for
nondeterminism.  If we increase the delay to 0.01 seconds and the
number of replays of a test used to check for nondeterminism, we
almost double the number of nondeterministic tests discovered.  Both
visible value and final state approaches find 40 nondeterministic
tests, though the overlap of tests thus detected is not quite perfect---each method detects two tests the other does not.  The difference
in overhead, interestingly, is very similar:  0.48\%; however, under
these parameters, visible value nondeterminism is actually cheaper on average
than final state nondeterminism, due to early termination when nondeterminism is detected ($p < 1.44 \times 10^{-10}$).

@@

We introduced a subtle bug into {\tt pyfakefs}, where the {\tt remove}
call checks that its target is not a directory, and returns the
correct error, but still carries out the remove.  Using {\tt
  os.remove} to delete directories violates the Python {\tt os}
specification (and the POSIX standard).  Detecting this bug using the TSTL
{\tt pyfakefs} harness is normally impossible without using another
file system as a reference.  However, the fault was detected  essentially
immediately with our failure determinism checks (an affirmative answer
to {\bf RQ1}).  Moreover, the overhead for the check in a
version of the code without the error was
less than 8\%  ({\bf RQ2}).  Detecting the fault using a reference file system 
required 17\% more testing time before detection, and took over twice as
long to reduce the failure, to a
slightly longer failing test, which did not have {\tt remove} as its
final operation (since further operations are required to 
expose the bad state).
Because vertical
reduction does not require running complete tests multiple times, and
does not affect the delta-debugging algorithm's performance, reducing
the failing test to 3 steps required less than a second ({\bf RQ3}).
Here, failure determinism even improves on hand-tuned differential testing.

@@

Non-differential testing, without failure determinism checks,
consistently (in all runs) killed
872 mutants, a 37.1\% kill rate.
%Adding a failure determinism check
%allowed the testing to (again, consistently) kill an additional 98 mutants, an 11\% improvement.
The differential harness, with an additional 60 seconds of testing
time (to make sure we accounted for the observed 17\%
extra time to detect in the manually constructed example: we want to
maximize the chance to detect a bug using the strong version of the harness), consistently killed 1,148 mutants, improving the kill rate to 48.8\%.  Failure
nondeterminism checking consistently added an
additional 71 mutants to that total, \emph{a 6\% improvement even for
this strong differential harness with a larger test budget}, nearly as
large an improvement as the gain from using a full reference
implementation for comparison.  Using
failure nondeterminism was the \emph{only} way to push the mutation score
above 50\%.  The kill rates are generally low
because {\tt universalmutator} includes many operations
thatproduce hard-to-kill
non-equivalent mutants not produced by other mutation tools.  This is a strong affirmative answer to {\bf
  RQ1}.

@@

There are several primary threats to validity:  first, our empirical
results are limited to a small set of Python programs, ranging from
relatively small and simple to large and complex libraries; the
representative nature of these subjects is not clear.
Because
no other tools implement the kind of approach taken here, we were unable
to perform a meaningful comparison with another nondeterminism
detection tool.  DeFlaker \cite{bell2018d}
performs a completely different task, and cannot check
nondeterminism at any smaller granularity than a whole test.  iDFlakies
\cite{idflakies}, similarly, ``does not further classify the causes of
the flaky tests'' beyond identifying them as likely related to test
order.  
Another threat to validity is posed by implementation errors.
%TSTL and the mutation tool are tested for bugs by a set of Travis CI
%tests, and the nondeterminism
%detected by TSTL can be demonstrated with standalone tests.

>>>>

Reducing a lengthy test showing nondeterminism of the swapsort usually takes less than two
minutes (out of 10 trials, only one took four minutes).  The resulting test will be usually be very compact, e.g.
reducing the test from 56 to 8 steps.

@@

Using the original TSTL harness, just over 20\% of all {\tt redis-py}
regression tests (of length 200) generated were flaky.  We determined
that the mean time to find all (known to us, over many experimental runs)
sources of nondeterminism and produce minimal tests is 3.2 
hours.  Our answer to {\bf RQ1} is thus a strong affirmative.

\begin{comment}
  Using the
nondeterminism checker to reduce flaky tests to a minimal set allows
us to set up an overnight run that will, in 12 hours, produce
a set of minimal nondeterministic tests exposing the root sources of
nondeterminism/flakiness: we set the harness to repeatedly run, then
produce a regression test.  If the test is flaky (exhibits
nondeterminism in result, over a set of 3 runs), we use the TSTL reducer to shrink it to a
normaliized \cite{onetest} test.  At the end of the testing period,
all distinct tests exhibiting nondeterminism are reported to the
user.  The process reduces the large number of different tests
exhibiting flakiness to a small set of tests, each short, showing a
different cause of nondeterminism.  
There is some duplication of causes, but the tests are sufficient for
identifying all nondeterminism sources.

This process is unlikely to truly require 12
hours, of course; our point is that it is completely automated and
can be given time proportional to the need to find even
low-probability sources of nondeterminism.  The actual amount of time given
may be smaller, or, for that matter, larger (leaving an important
library to self-document nondeterminism over a weekend is hardly a bad
practice).  In this case, the probability of flakiness
is not important, just the possibility, so we let the reducer
create a minimal test with some non-zero (but possibly very small)
observed probability of nondeterminism.  Our 12 hour run identified
all sources of nondeterminism successfully, according to a series of
48 hour runs performed on the version of the harness we modified to
remove nondeterminism, which produced no flaky tests.
\end{comment}

@@

We removed 11 calls from the {\tt redis-py} harness, and  modified 2
calls, after
identification of sources of nondeterminism.  After making these changes, no flakiness was observed in a sample of over 2,000 length
200 tests.   Resulting code coverage loss was minimal.
Mean branch coverage  was reduced
by less than 5 branches (a less than 1\% decrease).  AFL stability was 56.26\% with the harness allowing 
nondeterminism, but rose to 98.32\% using the harness with 
nondeterministic operations removed.

The overhead for
determinism checks, with no delay between operations, is only
20\%, much lower than the expected cost of running each test twice,
answering {\bf RQ2} also in the affirmative.

As to {\bf RQ3}, Figures \ref{fig:r1}-\ref{fig:r3} show delta-debugging of a typical
{\tt redis-py} nondeterministic test, originally with 287 steps.  The initial test behaves
nondeterministically about half the time.  Unmodified delta-debugging
eventually produces a test of length 16 that only behaves
nondeterministically 24\% of the time.  The reduction takes only 38
seconds.  For the purpose of identifying commands leading to
nondeterminism, this is acceptable.  However, if we were actually
debugging a complex nondeterminism bug in Redis itself, we might want
a more reliably nondeterministic test.  Using the same parameters as
in Section \ref{sec:pubbias}, we see the same pattern.  Simply making
a predicate that ``forces'' the probability to remain high, with a
large number of samples, does not work (Figure \ref{fig:r3}) and
requires over 2,000 seconds to produce a test with an even worse
probability of nondeterminism.  Using 10 samples and 10 replications,
on the other hand, actually improves the probability of
nondeterministic behavior, and (in this case) even produces a smaller test (only 14 steps) in just
over 10 minutes.

\begin{comment}
\subsubsection{Visible Value vs. Final State Nondeterminism}

We also used {\tt redis-py} to investigate the tradeoff between
overhead and ability to detect nondeterminism when using the two
proposed approaches to horizontal determinism.  We produced 300 tests
of length 100, using the known-nondeterministic version of the {\tt
  redis-py} harness.  We then used {\tt nondeterministic} and {\tt
  stepNondeterministic} calls, both with a delay of 0.005 seconds and
a single additonal execution of the test, to
check these tests for nondeterministic behavior.  Final state
nondeterminism actually detected one instance (but not root cause) of nondeterminism that
visible value nondeterminism failed to detect (21 detections
vs. 20). In general, final state
nondeterminism is less able to detect nondeterminism, but in this
instance the difference is less important than the nondeterminism of
whether a particular test will exhibit nondeterministic behavior in a
single run.  This difference is obviously not statistically
significant; for the nondeterminism in {\tt redis-py}, with these
parameters, the two approaches cannot be statistically distinguished
as to effectiveness in detecting nondeterminism.
Final state nondeterminism was also faster; visible value
nondeterminism took, on average, 0.48\% longer to check for
nondeterminism on the 300 tests, 1.199 mean seconds vs. 1.205 mean seconds.  The difference was significant, with
$p < 1.19 \times 10^{-23}$.

%\begin{comment}
These results, of course, depend on the parameters of the check for
nondeterminism.  If we increase the delay to 0.01 seconds and the
number of replays of a test used to check for nondeterminism, we
almost double the number of nondeterministic tests discovered.  Both
visible value and final state approaches find 40 nondeterministic
tests, though the overlap of tests thus detected is not quite perfect---each method detects two tests the other does not.  The difference
in overhead, interestingly, is very similar:  0.48\%; however, under
these parameters, visible value nondeterminism is actually cheaper on average
than final state nondeterminism, due to early termination when nondeterminism is detected ($p < 1.44 \times 10^{-10}$).

@@

We introduced a subtle bug into {\tt pyfakefs}, where the {\tt remove}
call checks that its target is not a directory, and returns the
correct error, but still carries out the remove.  Using {\tt
  os.remove} to delete directories violates the Python {\tt os}
specification (and the POSIX standard).  Detecting this bug using the TSTL
{\tt pyfakefs} harness is normally impossible without using another
file system as a reference.  However, the fault was detected  essentially
immediately with our failure determinism checks (an affirmative answer
to {\bf RQ1}).  Moreover, the overhead for the check in a
version of the code without the error was
less than 8\%  ({\bf RQ2}).  Detecting the fault using a reference file system 
required 17\% more testing time before detection, and took over twice as
long to reduce the failure, to a
slightly longer failing test, which did not have {\tt remove} as its
final operation (since further operations are required to 
expose the bad state).
Reducing
the failing test to just 3 steps required less than a second ({\bf RQ3}).

@@

Non-differential testing, without failure determinism checks,
consistently (in all runs) killed
872 mutants, a 37.1\% kill rate.
%Adding a failure determinism check
%allowed the testing to (again, consistently) kill an additional 98 mutants, an 11\% improvement.
The differential harness, with an additional 60 seconds of testing
time (to make sure we accounted for the observed 17\%
extra time to detect in the manually constructed example: we want to
maximize the chance to detect a bug using the strong version of the harness), consistently killed 1,148 mutants, improving the kill rate to 48.8\%.  Failure
nondeterminism checking consistently added an
additional 71 mutants to that total, \emph{a 6\% improvement even for
this strong differential harness with a larger test budget}, nearly as
large an improvement as the gain from using a full reference
implementation.  Using
failure nondeterminism was the \emph{only} way to push the mutation score
above 50\%.  The kill rates are generally low
because {\tt universalmutator} includes many operations
thatproduce hard-to-kill
non-equivalent mutants not produced by other mutation tools.  This is a strong affirmative answer to {\bf
  RQ1}.

@@

Our empirical
results are limited to a small set of Python programs, ranging from
relatively small and simple to large and complex libraries; the
representative nature of these subjects is not clear.
Because
no other tools implement the kind of approach taken here, we were unable
to perform a meaningful comparison with another nondeterminism
detection tool.  
%TSTL and the mutation tool are tested for bugs by a set of Travis CI
%tests, and the nondeterminism
%detected by TSTL can be demonstrated with standalone tests.
